---
title: "Maria Reyes Homework2 DACSS 690C"
output:
  html_document:
    df_print: paged
---

1.  Read from YOUR OWN GitHub (cloud): **dataGitLink** above should point to your repo (not mine)

```{r}
library(igraph)
library(ggraph)
library(tidygraph)
library(ggplot2)
library(GGally)
library(dplyr)

library(igraph)

dataGitLink <- "https://raw.githubusercontent.com/MariaReyesMedina/Homework-2---DACSS690C/main/hollywood.graphml"

hollywood <- read_graph(dataGitLink, format = "graphml")


```


```{r}

summary(hollywood)

vcount(hollywood)
ecount(hollywood)

```
2.  Compute node central measures: eigen_centrality, closeness, betweenness.

```{r}
V(hollywood)$eigen_centrality <-
  eigen_centrality(hollywood)$vector

V(hollywood)$closeness <-
  closeness(hollywood, normalized = TRUE)

V(hollywood)$betweenness <-
  betweenness(hollywood, normalized = TRUE)
```

3.  Prepare a correlation plot with previous measures, and **comment** the result.
```{r}
centrality_df <- data.frame(
    Actor = V(hollywood)$name,
    Eigen = V(hollywood)$eigen_centrality,
    Closeness = V(hollywood)$closeness,
    Betweenness = V(hollywood)$betweenness
  )

head(centrality_df)

ggpairs(
  centrality_df[,2:4]
)
```
The correlation plot shows that the three centrality measures are related in different ways. Closeness and Betweenness have a strong positive correlation (r = 0.766), indicating that actors who can reach others more efficiently also tend to act as bridges in the network. In contrast, Eigenvector centrality is moderately and negatively correlated with Closeness (r = -0.684), suggesting that actors connected to influential neighbors are not necessarily the ones closest to all other actors. Eigenvector and Betweenness show a weak negative relationship (r = -0.245), indicating that these measures capture different aspects of centrality.

4.  Redo the plot `finalGraph`, just showing the Hub nodes.

```{r}
HubNodes <- dplyr::slice_max(
  centrality_df,
  order_by = Eigen,
  n = 2
)$Actor

HubNodes


finalGraph <- ggraph(graph = hollywood) +
  geom_edge_link(
    aes(edge_width = as.ordered(weight)),
    color = "steelblue"
  ) +
  geom_node_point(
    aes(size = name %in% HubNodes)
  ) +
  geom_node_text(
    aes(
      label = ifelse(name %in% HubNodes, name, "")
    ),
    size = 5,
    nudge_y = 0.3,
    vjust = 0,
    color = "navy"
  ) +
  theme_graph()

finalGraph

```
5.  Answer: Is it possible communities are present? Support your answer with computations.

```{r}
louvain <- cluster_louvain(hollywood)
length(louvain)
modularity(louvain)
sizes(louvain)
```
The Louvain algorithm identified 3 communities with a modularity score of 0.412, suggesting a moderate community structure within the network.


6.  IF the presence of communities is possible, compute the communities using `Girvan-Newman` (aka edge_betweenness), and `Louvain` algorithms, or other algorithms. Plot only the best result

```{r}
girvan <- cluster_edge_betweenness(hollywood)
length(girvan)
modularity(girvan)
sizes(girvan)

```

```{r}

finalCommunityGraph <- ggraph(hollywood, layout = "fr") +
  geom_edge_link(
    color = "grey70"
  ) +
  geom_node_point(
    aes(color = factor(membership(girvan))),
    size = 5
  ) +
  geom_node_text(
    aes(label = name),
    size = 4,
    nudge_y = 0.3
  ) +
  theme_graph()

finalCommunityGraph

```
Since the presence of communities was possible, I applied both Louvain and Girvan-Newman (edge betweenness) algorithms. Louvain identified three communities with a modularity score of 0.412, while Girvan-Newman identified two communities with a modularity score of 0.420. Because Girvan-Newman produced the highest modularity value, it was selected as the best community detection method and used for the final visualization.


---


